\demes: a Python package for pain-free specification of complex demography

### 2021-04-23 paper.tex
some edits
some cleanup, getting ready for editing...

\documentclass[11pt]{article}
\usepackage[round,comma,authoryear]{natbib}
\usepackage{fullpage}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage[newfloat]{minted}
\usepackage{tcolorbox}
\usepackage{etoolbox}
\BeforeBeginEnvironment{minted}{\begin{tcolorbox}}
\AfterEndEnvironment{minted}{\end{tcolorbox}}
\usepackage{url}
\usepackage{fancyvrb}
\usepackage{caption}
\newenvironment{code}{\captionsetup{type=listing}\centering}{}
\SetupFloatingEnvironment{listing}{name=Demographic Model}

@@

The ability to simulate complex demographic scenarios, however, does not come
without its own set of obstacles and frustrations. First of all, specifying models
with many populations and parameters can be cumbersome and error-prone
(e.g., \citep{ragsdale2020lessons}). It is
time-consuming and tedious to implement and then verify the correctness
of such demographic models. Making matters worse, each software package typically has
its own application programming interface (API), which makes it difficult to translate
models between simulation methods.

Secondly, many software API are not particularly easy to read, especially when there
are a large number of demographic events and parameters. This not only makes
debugging difficult, but it can be difficult to even notice if a mistake has been
made. A common interface for unambiguously specifying demographic
models would reduce implementation errors and promote ease-of-use for many simulation
software packages.

@@

Here, we present \demes, a Python package for specifying complex demographic models
in a way that is both human readable and can be directly passed to a growing
number of simulation engines. \Demes\ supports model specification in a high-level
\YAML\ format \citep{ben2009yaml}
that is designed to be as easy as possible to implement, read, and
share. Models are then represented in an unambiguous low-level format within Python
that checks model validity and can then be used as the input for simulation software.
The initial release of \demes\ (version 1.0, described here) supports \emph{static}
demographic descriptions, that is, models with fixed parameters. Future releases
will support [drawing parameters from distributions, for example for ABC-based
inference, or describing ``dynamic`` models where parameters may be fit to data,
e.g. \moments\ or \dadi].

@@

\begin{code}
\captionof{listing}{An IM model.}
\begin{tcolorbox}
\inputminted[linenos,numbersep=5pt]{yaml}{models/IM.yaml}
\end{tcolorbox}
\label{code:im_model}
\end{code}

>>>>

\documentclass[11pt]{article}
\usepackage[round,comma,authoryear]{natbib}
\usepackage{fullpage}
\usepackage{authblk}
\usepackage{graphicx}

\usepackage[newfloat]{minted}
\usepackage{tcolorbox}
\usepackage{etoolbox}
\BeforeBeginEnvironment{minted}{\begin{tcolorbox}}
\AfterEndEnvironment{minted}{\end{tcolorbox}}
\usepackage{url}
\usepackage{fancyvrb}
\usepackage{caption}
\newenvironment{code}{\captionsetup{type=listing}\centering}{}
\SetupFloatingEnvironment{listing}{name=Demographic Model}
% can either input code directly in a minted environment
% or input from a source file using \inputminted

@@

The ability to simulate complex demographic scenarios, however, does not come
without its own set of obstacles and frustrations. First, specifying models
with many populations and parameters can be cumbersome and error-prone (e.g.,
\citep{ragsdale2020lessons}). It is time-consuming and tedious to implement and
then verify the correctness of such demographic models. Making matters worse,
each software package typically has its own application programming interface
(API), which makes it difficult to translate models between simulation
software.

Second, many software API are not particularly easy to read, especially when
there are a large number of demographic events and parameters. This not only
makes debugging difficult, but it can be difficult to even notice if a mistake
has been made. A common interface for unambiguously specifying demographic
models would reduce implementation errors and promote ease-of-use for many
simulation software packages.

@@

Here, we present \demes, a Python package for specifying complex demographic
models in a way that is both human readable and can be directly passed to a
growing number of simulation engines. \Demes\ supports model specification in a
high-level \YAML\ format \citep{ben2009yaml} that is designed to be as easy as
possible to implement, read, and share. Models are then represented in an
unambiguous low-level format within Python that checks model validity and can
then be used as the input for simulation software.  The initial release of
\demes\ (version 0.1, described here) supports \emph{static} demographic
descriptions, that is, models with fixed parameters. Future releases will
support [drawing parameters from distributions, for example for ABC-based
inference, or describing ``dynamic`` models where parameters may be fit to
data, e.g. \moments\ or \dadi].

@@

\begin{code}
\begin{tcolorbox}
\inputminted[linenos,numbersep=5pt]{yaml}{models/IM.yaml}
\end{tcolorbox}
\captionof{listing}{An IM model.}
\label{code:im_model}
\end{code}
